Deriving Multi-Headed Planar Dependency Parses from Link Grammar Parses

### 2014-12-13 doc/slides/hong+eisner.TLT13.slides.tex
url link citation for SCIP

\frame{\frametitle{Link Grammars}
  \begin{itemize}
  \item[] Grammar-based formalism for projective dependency parsing with undirected links
  \item[] Orignal formalism and English Link Grammar created by Davy Temperley, Daniel Sleator, John Lafferty (1991)
  \end{itemize}
}

@@

\subsection{What is ILP?}
\frame{\frametitle{What is Integer Linear Programming?}
\begin{itemize}
  \item Its an optimization problem where some or all of the variables are integers. 
  \item The objective function and constraints are linear.
  \item Its NP-Hard! But good solvers exist that work well most of the time.
  \item Our ILP is encoded as a ZIMPL program and solved using the SCIP Optimization Suite
\end{itemize}
}

@@

\frame{\frametitle{Link Results: Subject-Verb links are backwards}
\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
	\end{deptext}
	\deproot[edge below, edge style = {blue, dotted}]{3}{\small ROOT}
	\depedge[edge above, edge style = {red, ultra thick}]{2}{3}{S}
	\depedge[edge below, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
	\depedge[edge below, edge style = {blue, dotted}]{3}{4}{\small ADV}
	\depedge[edge below, edge style = {blue, dotted}]{3}{5}{\small ADV}
	\depedge[edge above, edge style = {red, thick}]{3}{6}{I}
	\depedge[edge below, edge style = {blue, thick}]{3}{6}{\small VC}
	\depedge[edge below, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
	\depedge[edge above, edge style = {red, thick}]{2}{1}{D}
	\depedge[edge below, edge style = {blue, thick}]{2}{1}{\small NMOD}
	\depedge[edge above, edge style = {red, thick}]{7}{8}{MV}
	\depedge[edge below, edge style = {blue, thick}]{7}{8}{\small ADV}
	\depedge[edge above, edge style = {orange, thick}]{6}{7}{P}
	\depedge[edge below, edge style = {blue, thick}]{6}{7}{\small VC}
	\depedge[edge above, edge style = {red, thick}]{8}{9}{J}
	\depedge[edge below, edge style = {blue, thick}]{8}{9}{\small PMOD}
	\deproot[edge above, edge style = {red, dotted}]{2}{W}
	\deproot[edge above, edge style = {orange, ultra thick, dotted}]{7}{WV}
	\deproot[edge above, edge style = {red, dotted}]{10}{X}
	\depedge[edge above, edge style = {red, dotted}]{6}{4}{E}
	\depedge[edge above, edge style = {red, dotted}]{6}{5}{E}
  \end{dependency}
\end{figure}

@@

\frame{\frametitle{Link Results: Subject-Verb links are backwards}
\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
	\end{deptext}
	\deproot[edge below, edge style = {gray}, hide label]{3}{\small ROOT}
	\depedge[edge above, edge style = {red, ultra thick}]{2}{3}{S}
	\depedge[edge below, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
	\depedge[edge below, edge style = {gray}, hide label]{3}{4}{\small ADV}
	\depedge[edge below, edge style = {gray}, hide label]{3}{5}{\small ADV}
	\depedge[edge above, edge style = {gray}, hide label]{3}{6}{I}
	\depedge[edge below, edge style = {gray}, hide label]{3}{6}{\small VC}
	\depedge[edge below, , edge style = {gray}, hide label, edge unit distance =1.5ex]{3}{10}{\small P}
	\depedge[edge above, edge style = {gray}, hide label]{2}{1}{D}
	\depedge[edge below, edge style = {gray}, hide label]{2}{1}{\small NMOD}
	\depedge[edge above, edge style = {gray}, hide label]{7}{8}{MV}
    \depedge[edge below, edge style = {gray}, hide label]{7}{8}{\small ADV}
	\depedge[edge above, edge style = {gray}, hide label]{6}{7}{P}
	\depedge[edge below, edge style = {gray}, hide label]{6}{7}{\small VC}
	\depedge[edge above, edge style = {gray}, hide label]{8}{9}{J}
	\depedge[edge below, edge style = {gray}, hide label]{8}{9}{\small PMOD}
	\deproot[edge above, edge style = {gray}, hide label]{2}{W}
    \deproot[edge above, edge style = {gray}, hide label]{7}{WV}
	\deproot[edge above, edge style = {gray}, hide label]{10}{X}
	\depedge[edge above, edge style = {gray}, hide label]{6}{4}{E}
	\depedge[edge above, edge style = {gray}, hide label]{6}{5}{E}
  \end{dependency}
\end{figure}

@@

\frame{\frametitle{Link Results: Subject-Verb links are backwards}

@@

\frame{\frametitle{Link Results: Subject-Verb links are backwards}
\begin{itemize}
      %\item Different dependency formalisms disagree on whether a clause is headed by the aux verb or the main verb.
  \uncover<1->{\item The Link Grammar seems to be inconsistent about whether the auxiliary verb or the main verb is the head of a clause. }
  \uncover<2->{\item Sometimes the governing verb links to the auxilliary, and sometimes to the main, depending on the type of clause. }
  \uncover<3->{\item But the governing verb usually links to the subject when there is one.}
  \uncover<4->{\item So this makes the subject a consistent choice to make the head of a clause.}
\end{itemize}

@@

\section{Conclusions}
\frame{\frametitle{Conclusions}
  \begin{itemize}
  \item Link Grammar parses can be oriented into connected DAGs
  \item New corpora for building multi-headed dependency parsers
  \item ILP can be used to build or annotate corpora
  \end{itemize}
}

>>>>

\frame{\frametitle{Link Grammars}
  \begin{itemize}
  \item[] Grammar-based formalism for projective dependency parsing with undirected links
  \item[] Orignal formalism and English Link Grammar created by Davy Temperley, Daniel Sleator, and John Lafferty (1991)
  \end{itemize}
}

@@

\subsection{What is ILP?}
\frame{\frametitle{What is Integer Linear Programming?}
\begin{itemize}
  \item Its an optimization problem where some or all of the variables are integers. 
  \item The objective function and constraints are linear.
  \item Its NP-Hard! But good solvers exist that work well most of the time.
  \item Our ILP is encoded as a ZIMPL program and solved using the SCIP Optimization Suite\footnote{\url{http://scip.zib.de/}} 
\end{itemize}
}

@@

\subsection{Subject-Verb Links}
\frame{\frametitle{Link Results: Subject-Verb Links are Backwards}
\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
	\end{deptext}
	\deproot[edge below, edge style = {blue, dotted}]{3}{\small ROOT}
	\depedge[edge above, edge style = {red, ultra thick}]{2}{3}{S}
	\depedge[edge below, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
	\depedge[edge below, edge style = {blue, dotted}]{3}{4}{\small ADV}
	\depedge[edge below, edge style = {blue, dotted}]{3}{5}{\small ADV}
	\depedge[edge above, edge style = {red, thick}]{3}{6}{I}
	\depedge[edge below, edge style = {blue, thick}]{3}{6}{\small VC}
	\depedge[edge below, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
	\depedge[edge above, edge style = {red, thick}]{2}{1}{D}
	\depedge[edge below, edge style = {blue, thick}]{2}{1}{\small NMOD}
	\depedge[edge above, edge style = {red, thick}]{7}{8}{MV}
	\depedge[edge below, edge style = {blue, thick}]{7}{8}{\small ADV}
	\depedge[edge above, edge style = {orange, thick}]{6}{7}{P}
	\depedge[edge below, edge style = {blue, thick}]{6}{7}{\small VC}
	\depedge[edge above, edge style = {red, thick}]{8}{9}{J}
	\depedge[edge below, edge style = {blue, thick}]{8}{9}{\small PMOD}
	\deproot[edge above, edge style = {red, dotted}]{2}{W}
	\deproot[edge above, edge style = {orange, ultra thick, dotted}]{7}{WV}
	\deproot[edge above, edge style = {red, dotted}]{10}{X}
	\depedge[edge above, edge style = {red, dotted}]{6}{4}{E}
	\depedge[edge above, edge style = {red, dotted}]{6}{5}{E}
  \end{dependency}
\end{figure}

@@

\frame{\frametitle{Link Results: Subject-Verb Links are Backwards}
\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
	\end{deptext}
	\deproot[edge below, edge style = {gray}, hide label]{3}{\small ROOT}
	\depedge[edge above, edge style = {red, ultra thick}]{2}{3}{S}
	\depedge[edge below, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
	\depedge[edge below, edge style = {gray}, hide label]{3}{4}{\small ADV}
	\depedge[edge below, edge style = {gray}, hide label]{3}{5}{\small ADV}
	\depedge[edge above, edge style = {gray}, hide label]{3}{6}{I}
	\depedge[edge below, edge style = {gray}, hide label]{3}{6}{\small VC}
	\depedge[edge below, , edge style = {gray}, hide label, edge unit distance =1.5ex]{3}{10}{\small P}
	\depedge[edge above, edge style = {gray}, hide label]{2}{1}{D}
	\depedge[edge below, edge style = {gray}, hide label]{2}{1}{\small NMOD}
	\depedge[edge above, edge style = {gray}, hide label]{7}{8}{MV}
    \depedge[edge below, edge style = {gray}, hide label]{7}{8}{\small ADV}
	\depedge[edge above, edge style = {gray}, hide label]{6}{7}{P}
	\depedge[edge below, edge style = {gray}, hide label]{6}{7}{\small VC}
	\depedge[edge above, edge style = {gray}, hide label]{8}{9}{J}
	\depedge[edge below, edge style = {gray}, hide label]{8}{9}{\small PMOD}
	\deproot[edge above, edge style = {gray}, hide label]{2}{W}
    \deproot[edge above, edge style = {gray}, hide label]{7}{WV}
	\deproot[edge above, edge style = {gray}, hide label]{10}{X}
	\depedge[edge above, edge style = {gray}, hide label]{6}{4}{E}
	\depedge[edge above, edge style = {gray}, hide label]{6}{5}{E}
  \end{dependency}
\end{figure}

@@

\frame{\frametitle{Link Results: Subject-Verb Links are Backwards}

@@

\frame{\frametitle{Link Results: Subject-Verb Links are Backwards}
\begin{itemize}
      %\item Different dependency formalisms disagree on whether a clause is headed by the aux verb or the main verb.
  \uncover<1->{\item The Link Grammar seems to be inconsistent about whether the auxiliary verb or the main verb is the head of a clause. }
  \uncover<2->{\item Sometimes the governing verb links to the auxilliary, and sometimes to the main, depending on the type of clause. }
  \uncover<3->{\item But the governing verb usually links to the subject when there is one.}
  \uncover<4->{\item So this makes the subject a consistent choice to make the head of a clause.}
\end{itemize}

@@

\section{Conclusions}
\frame{\frametitle{Conclusions}
  \begin{itemize}
  \item Link Grammar parses can be oriented into connected DAGs
  \item new corpora for building multi-headed dependency parsers
  \item ILP can be used to build or annotate corpora. 
  \end{itemize}
}
